---
title: "Introduction to the R package survstan"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to the R package survstan}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

To fit a survival model with the __survstan__ package, the user must choose among one of the available fitting functions `*reg()`, where * stands for the type of regression model, that is, `aft` for accelerated failure time (AFT) models, `ah` for accelerates hazards (AH) models, `ph` for proportional hazards (PO) models, `po` for proportional (PO) models, or `yp` for Yang \& Prentice (YP) models.

The specification of the survival formula passed to the chosen `*reg()` function follows the same syntax adopted in the __survival__ package, so that transition to the __survstan__ package can be smoothly as possible for those familiar with the __survival__ package.

The code below shows the model fitting of a AFT model with Weibull baseline distribution using the `survstan::aftreg()` function. For comparison purposes, we also fit to the same data the Weibull regression model using `survival::survreg()` function:

```{r setup, message=FALSE}
library(survstan)
library(dplyr)

ovarian <- ovarian %>%
  mutate(
    across(c("rx", "resid.ds"), as.factor)
  )

survreg <- survreg(
  Surv(futime, fustat) ~ ecog.ps + rx, 
  dist = "weibull", data = ovarian
)

survstan <- aftreg(
  Surv(futime, fustat) ~ ecog.ps + rx, 
  dist = "weibull", data = ovarian
)

```

Although the model specification is quite similar, there are some important differences that the user should be aware of. While the model fitted using the `survival::survreg()` function uses the log scale representation of the AFT model with the presence of an intercept term in the linear predictor, the `survstan::aftreg()` considers the original time scale for model fitting without the presence of the intercept term in the linear predictor.

To see that, let us summarize the fitted models:

```{r}
summary(survreg)
summary(survstan)
```

Next, we show how to fit a PH model using the `survstan::phreg()` function. For comparison purposes, the semiparametric Cox model is also fitted to the same data using the function `survival::coxph()`.

```{r}
survstan <- phreg(
  Surv(futime, fustat) ~ ecog.ps + rx, 
  data = ovarian, dist = "weibull"
)
cox <- coxph(
  Surv(futime, fustat) ~ ecog.ps + rx, data = ovarian
)
coef(survstan)
coef(cox)
```


<!-- Now we  and the semiparametric proportional hazards model using the `survival::coxph()` -->

<!-- The usege of other functions implemented in the __survstan__ package are illustrated in the code presented below: -->

<!-- ```{r} -->

<!-- # fitting the model: -->
<!-- fit <- aftreg( -->
<!--   Surv(futime, fustat) ~ .,  -->
<!--   dist = "weibull", data = ovarian -->
<!-- ) -->

<!-- # investigating the fitted model: -->
<!-- estimates(fit) -->
<!-- coef(fit) -->
<!-- confint(fit) -->
<!-- summary(fit) -->
<!-- tidy(fit) -->
<!-- vcov(fit) -->


<!-- # residual plots: -->
<!-- ggresiduals(fit, type = "coxsnell") -->
<!-- ggresiduals(fit, type = "martingale") -->
<!-- ggresiduals(fit, type = "deviance") -->


<!-- # Deviance analysis: -->
<!-- fit1 <- aftreg(Surv(futime, fustat) ~ 1, baseline = "weibull", data = ovarian, init = 0) -->
<!-- fit2 <- aftreg(Surv(futime, fustat) ~ rx, baseline = "weibull", data = ovarian, init = 0) -->
<!-- fit3 <- aftreg(Surv(futime, fustat) ~ rx + ecog.ps, baseline = "weibull", data = ovarian, init = 0) -->

<!-- anova(fit1, fit2, fit3) -->


<!-- # Model comparison using AIC: -->
<!-- AIC(fit1, fit2, fit3) -->

<!-- ``` -->

